---
title: "Hello Qoma Smuggler!"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,comment='')
```

The R package _qoma.smuggler_  provides high-level access to the FAME HLI.  To open the FAME HLI,
```{r}
library(ggplot2)
library(qoma.smuggler)

if(!open_hli())knitr::knit_exit()
```
To move FAME data across the border to R, provide a database name:
```{r}
dbname <- paste(Sys.getenv("FAME"),"util","driecon.db",sep="/")
cat(dbname)
```
Read the database into a _qoma.smuggler_ `List`:
```{r}
famedata <- read_fame(dbname)
```
To display all the contents as a catalog, 
```{r}
print_catalog(famedata,list.len=3)
```
You may display individual object's meta data from a _qoma.smuggler_ `List`. 
```{r}
cat(famedata$get_meta('GDP'))
```
FAME series are accessible as a `tibble` structure with dates expressed as _lubridate_ package structures `Date`:
```{r}
tbl<-famedata$get_data('GDP')
str(tbl)
```
The _tibble_ package is part of the _tidyverse_ ecosystem and works well with packages such as _ggplot_.
```{r}
theme_set(theme_minimal())
ggplot(data = tbl, aes(x = date, y = GDP))+
  geom_line(color = "#00AFBB", size = 2)
```

To move R data back across the border to FAME,
```{r}
mydb <- file.path(tempdir(),"mycopy.db")
write_fame(mydb,famedata)
```

When you complete your work, close the FAME HLI to avoid data loss,
```{r}
close_hli()
file.remove(mydb)
```