---
title: "Read FAME databases from R"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,comment='')
```

FAME databases map names to data structures.  FAME data types include:

BOOLEAN 
DATE (at various frequencies)
NUMERIC
PRECISION
STRING

The _rhli_ package maps FAME data types to compatible basic R data types. 
The _qoma.smuggler_ package constructs more complex _tidyverse_ data structures such as `Date` and `tibble`. _qoma.smuggler_ defines an R reference class `List` which implements a mutable `list` to hold FAME data.

The utility function `open_hli()` opens the FAME environment and prints diagnostic information.
```{r}
library(qoma.smuggler)
library(rhli)

if(!open_hli())knitr::knit_exit()
```

The FAME distribution includes a number of sample databases.  We will use the DRI Economics _driecon_ sample database for this notebook.
```{r}
dbname <- paste(Sys.getenv("FAME"),"util","driecon",sep="/")
cat(dbname)
```

The _qoma.smuggler_ function `read_fame()` reads FAME data objects into a _qoma.smuggler_ `List`.
By default, the function `read_fame()` loads the entire specified database to a  _qoma.smuggler_ `List`.
```{r}
famedata <- read_fame(dbname)
```
The method `get_data(objectName)` returns FAME data for objectName from the _qoma.smuggler_ `List`.
    For FAME SCALAR objects, `get_data(objectName)` returns one value. 
    For FAME SERIES objects, `get_data(objectName)` returns a `tibble` with multiple values.
```{r}    
famedata$get_data('GDP')
```
The method `get_meta(objectName)` returns FAME object meta data as a string from a  _qoma.smuggler_ `List`.

```{r}
cat(famedata$get_meta('GDP'))
```
Specifying a wildcard pattern will select a subset of the data objects with a compatible pattern.  You may  
specify an full name, retrieving one series.
```{r}
wilnam <- "ip?"
famedata <- read_fame(dbname,wilnam)
```
The function `print_catalog()` will display summary meta data for the contents retrieved. For brevity, 
we specify optional parameter `list.len = 3` to limit output.
```{r}
print_catalog(famedata,list.len = 3)
```
Specifying a date range will select a subset of the database with specified frequency.  Here we use the FAME HLI
constant `HANDEC` which means _ANNUAL(DECEMBER)_ frequency. See the FAME support website CHLI documentation for available codes.
Data returned will be limited to the specified date range.
```{r}
range <- to_fame_range(HANDEC,"1993","2002")
famedata <- read_fame(dbname,fame_range_ = range)
print_catalog(famedata,list.len = 2)
```
Note the data retrieved is for the specified date range only; not an objects full date range:
```{r}
famedata$get_data('GDP')
```
The _qoma.smuggler_ utility function `close_hli()` closes the FAME environment.
```{r}
close_hli()
```