---
title: "Write to a FAME database from R"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment='')
```

```{r}
library(qoma.smuggler)
library(rhli)

if(!open_hli())knitr::knit_exit()
```

Use FAME HLI monthly frequency constant `HMONTH` and FAME date literals "18m1" and "18m2" --- 
abbreviations for the first and twelfth month of the year 2018.
```{r}
rng <- to_fame_range(HMONTH,"18m1","18m12")
rng
```
Convert a FAME date range `rng` to a _lubridate_ index.
```{r}
tbl <- to_lubridate_index(rng)
tbl
```
Generate some normal variates in R, add to the tibble as column 'x'.
```{r}
nobs <- rng[3]-rng[2]+1
tbl['x'] <- rnorm(nobs)
tbl
```
Setup a _qoma.smuggler_ `List` structure to hold the data.  

See what it looks like with the `print_catalog()` 
function.
```{r}
mydb <- List() # mutable list
entry <- newEntry( # entry with data and FAME metadata
  tbl$x,
  desc = "N(0,1)",
  docu = "R generated N(0,1) time series.",
  range = rng,
  obse = rhli::HOBSUM
  )
mydb$put('x',entry)
print_catalog(mydb)
mydb$get('x')
```
Write the contents of the _qoma.smuggler_ `List` to a FAME database.
```{r}
dbfile <- file.path(tempdir(),"tmp.db")
write_fame(dbfile,mydb)
```
Use 4GL to peek at data in FAME.
```{r}
txtfile <- file.path(tempdir(),"tmp.txt")
cmd <- rhli::Character(paste(
  "open<acc read>\"",dbfile,"\" as db;",  
  "output<acc over>\"",txtfile,"\";",
  "cata db;",  
  "whats x;",  
  "disp x;",  
  "output terminal;",
  "close db;",
  sep=""))
rhli::cfmfame(rhli::Integer(-1), cmd)
cat(readLines(txtfile), sep = '\n')
```
```{r}
close_hli()

file.remove(dbfile)
file.remove(txtfile)
```
